|

### 2021-12-02 04-Chp-Systematic-Review-Results.Rmd
End of day:

---
bibliography: bibliography/references.bib
csl: bibliography/nature.csl
output:
  bookdown::pdf_document2:
    template: templates/brief_template.tex
  bookdown::word_document2: 
      toc: false
      toc_depth: 3
      reference_docx: templates/word-styles-reference-01.docx
      number_sections: false 
  bookdown::html_document2: default
documentclass: book
---
```{block type='savequote', include=knitr::is_latex_output(),quote_author='(ref:sys-rev-results-quote)', echo = TRUE}
Hold  
```

@@

## Lay summary

Using the methods outlined in the previous chapter, this chapter presents a systematic review of primary studies that have examined the relationship between the levels of blood lipids (such as cholesterol and triglycerides) and dementia outcomes. In addition, it examined the relationship betweeen treatments that change blood lipid levels, such as statins, and dementia outcomes

My review included `r n_included` primary studies that contained information on this relationship. I found that statins appear to reduce the risk of Alzheimer's disease, but had no effect of vascular dementia. Lipid levels were not associated with any outcome. The methods used in some of the primary studies meant that I was less confident in their results.

The added value of including preprinted study reports, made possible using the tool described in the Chapter \@ref(sys-rev-tools-heading), along with the use of the results of this review in subsequent chapters, is discussed.

@@

## Results

@@

The database search identified `r num_to_text(23447)` records, of which `r comma(7338)` were duplicated records. Of the random sample of 500 records screened to ensure the accuracy of the study design filters, no eligible records were identified. Many of those excluded by the filters were commentaries/educational articles, or described basic science studies. <!-- TODO include some example citations here -->

@@

The PRISMA flow diagram <!-- TODO CITATION NEEDED - cite PRIMSA flow paper here, if published --> presented in Figure \@ref(fig:prisma-flow-fig), illustrates the movement of articles through the review. To highlight the contribution of preprint archives to the review, the flow diagram delineates between those records captured through databases searches (presented on the left of the diagram) and those captured by the search tool described in the previous chapter (presented in grey on the right of the diagram).

Common reasons for exclusion at the full text-stage included studies that reported on ineligible exposures (n = `r n_wrong_exposure` ; most commonly an ineligible lipid fraction) or outcomes (n=24; e.g. change in cognitive scores) or used an ineligible study design (n= `r n_study_design`).

@@

```{r prisma-flow-fig, echo = FALSE, results="asis", fig.pos="H", fig.cap='(ref:prisma-flow-cap)', out.width='100%', fig.scap='(ref:prisma-flow-scap)'}
knitr::include_graphics(file.path("figures/sys-rev/prismaflow.png"))
```

@@

For the assessment of the intra-/inter-rater reliability, the estimated values of $AC1$ were interpreted against the categories presented in Table \@ref(tab:gwet-table). For the inter-rater reliability, agreement was "almost perfect" ($AC1$ = `r agreeInter_coeff[1]`, $kappa$ = `r agreeInter_coeff[2]`, Table \@ref(tab:agreeInter-table)). Similarly for intra-rater reliability, agreement was "almost perfect" ($AC1$ = `r agreeIntra_coeff[1]`, $kappa$ = `r agreeIntra_coeff[2]`, Table \@ref(tab:agreeIntra-table)). The discrepancy between the $AC1$ and $kappa$ coefficients illustrates the sensitivity of $kappa$ to imbalanced marginals, caused in this sample by a large imbalance  towards exclusion.[@feinstein1990]

@@

Those records which were excluded in the initial screening, but were included by the second reviewer (n=`r discrepancy_Inter`, Table \@ref(tab:agreeInter-table)) were investigated. This discrepancy between the two reviewers was explained in all cases by differing interpretations of the inclusion criteria, most commonly around the definition of cognitive decline versus dementia and the definition of eligible lipids fractions.

@@

Following full-text screening, `r n_included` unique studies (described across `r n_reports_included` reports) met the criteria for inclusion in the review.`r get_all_citations()` Table \@ref(tab:studyCharacteristics-table) presents a summary of the characteristics of each study.

The majority of included studies described non-randomised analyses, with the sole two included randomised controlled trials (the Heart Protection Study/British Heart Foundation trial[@heartprotectionstudycollaborativegroup2002] and the JUPITER trial[@ridker2008]) both reporting on the effect of statin use on all-cause dementia in older adults. A similarly small number of Mendelian randomisation studies were identified, several of which employed a two-sample approach using summary statistics from the same published genome wide association studies (GWAS) leading to complications in the synthesis (see Section \@ref(sys-rev-res-AD)).

@@

<!-- TODO Enrure all abbreviations at the end of this table are in the abbreviation section at the start. -->

@@

As discussed above, the risk-of-bias assessments are presented alongside their corresponding numerical result. A more detailed discussion of the sources and directions of bias is presented in Chapter \@ref(tri-heading), and so this section presents a brief summary of the biases observed in each study design.

For the two randomised controlled trials, both were judged to be at low risk of bias. In contrast, many of the non-randomised studies of statin use were at serious risk of bias due poor controlling for confounding, immortal time bias, and missing outcome data (as noted above, a judgement of low risk of bias for a non-randomised study is rare). Similarly, non-randomised studies of exposures suffered from incomplete adjustment for potentially important confounders, and concerns over the selection of the reported result from among several analyses (e.g. examination of lipids as a binary or continuous variable). Finally, bias was introduced into Mendelian randomisation studies via the potential for hoziontal pleiotropy and population stratification.[@davies2018]

Following best practice, any result judged to be at critical risk of bias should be excluded from any quantitative analyses. <!-- TODO CITATION NEEDED --> `r num_to_text(n_critical,T)` observational studies were excluded on this basis, predominantly due to a lack of adjustment for any potentially important confounders (i.e. the study reported unadjusted estimates).`r critical_citations`

@@

(ref:statinsRCT-cap) Random-effects meta-analysis of randomised controlled trials examining statin statins on all-cause dementia

@@

<!-- TODO The number of studies in the inline text refer to the number of studies rather than the number of results. Note here that Tynkkyen examine multiple cohorts, which were extracted individually rather than as part of the reported meta-analysis. -->

@@

<!----------------------------------------------------------------------------->
(ref:lipidFractionsDementia-cap) Random-effects meta-analysis of four lipid fractions (total cholesteol, HDL, LDL, and triglycerides) on all-cause dementia risk, standardised per 1-SD increase in the lipid fraction.

@@

`r obsHyperchol$AD$n` studies reported on the association of hypercholesterolemia with all-cause dementia and provided weak evidence for an effect (`r obsHyperchol$AD$estimate`, Figure \@ref(fig:obsHyperAD))

@@

<!----------------------------------------------------------------------------->
(ref:lipidFractionsAD-cap) Random-effects meta-analysis of LDL on vascular dementia risk, standardised per 1-SD increase in the lipid fraction.

@@

Finally, there were several identified Mendelian randomisation studies examining the effect of genetically lowered LDL-c on Alzheimer's disease risk (Figure \@ref(fig:mrDuplication)). However, all of these studies used a two-sample approach, making use of summary statistics from the GLGC and IGAP consortia. Due to this overlap, which would result in a falsely precise estimate caused by multiple counting of the same participants, no meta-analysis of these studies was performed. However, across all analysis, using varying number of SNPS, no evidence for an effect was observed (Figure \@ref(fig:mrDuplication)).

@@

A single Mendelian randomisation analysis was identified that examining the effect of genetically lowered LDL-c levels via HMGCR inhibition on the risk of vascular dementia, which provided weak evidence for an effect (`r mrStatin$VaD$estimate`).

@@

Few studies investigated the effect of individual lipid fractions. Only for total cholesterol was there greater than one result reported (Figure \@ref(fig:lipidFractionsVaD)), and a meta-analysis of these found weak evidence for an effect (`r n_effect(obsLipids$VaD_TC)`). A single study provided evidence on the other three fractions,[@yoshitake1995] and similar found minimal evidence of an effect LDL-c (`r n_effect(obsLipids$VaD_LDL)`), HDL-c (`r n_effect(obsLipids$VaD_HDL)`) or triglycerides (`r n_effect(obsLipids$VaD_TG)`).

@@

@@

<!-- TODO Use the example of getting the cut-off points from the authors as an illustration of how reporting sucks -->

@@

There were `r n_dr_initial` studies that were eligible for dose-response meta-analysis as they provide several categories of lipid exposure. However, following data extraction, `r num_to_text(n_dr_excluded)` were excluded as they did not reported the relevant information needed (most commonly, the cut-off measures for each category).

Across the remaining `r num_to_text(n_dr_final)` studies, a sufficient number of studies (n $\geqslant$ 3) were identified only for the total cholesterol-Alzheimer's, LDL-Alzheimer's, and total cholesterol-dementia strata. This analysis provided weak evidence for a non-linear effect of lipid levels on dementia outcomes, and Figure \@ref(fig:lipidsDoseResponse) illustrates this for the total cholesterol-Alzheimer's strata. Similar figures for the other analysed lipid-outcome strata are presented in Appendix \@ref(hold).

@@

### Additional analyses

Investigation of potential sources of heterogeneity was complicated by two factors. In the first instance, few meta-analysis included more than 10 results, the recommended minimum required for meta-regression. Secondly, poor reporting of summary statistics including education level and baseline cognitive ability precluded the use of several results in a meta-regression analysis.

@@

<!-- TODO [_Note: Julian, is this sufficient or should I report statistics for each? Additionally, are there any other sensitivity analyses I should be presenting here?_] -->

<!-- One particularly interesting meta-bias potentially applicable in this review is the definition of code lists across -->

<!-- Many studies using large observational electronic health records (n=?) did not report the code-lists used define the outcome events in their analysis. Depsite attempts to obtain this additional information through contact with a -->

### Missing evidence

@@

As show in Figure \@ref(fig:prisma-flow-fig), the number of hits returned by the preprint searching was not substantial (bioRxiv = 256, medRxiv = 0). From these hits, three preprinted reports of eligible studies were included in the review, of which two described unique studies not captured by the main search.[@andrews2019; @so2017]

One preprint (So _et al_[@so2017]) provided additional evidential value in a single meta-analysis (`r comma(1/18*100)`% of all meta-analyses performed in this chapter). This meta-analysis of Mendelian randomisation studies examined the effects of lipid-lowering via HMGCR mutations on incidence of Alzheimer's disease (Figure \@ref(fig:mrStatinADFig). To assess the evidence added to the meta-analysis through inclusion of the preprint, the data were re-analysed using a fixed-effect model. Examination of the weight assigned to each result in the analysis illustrates that a large proportion (`r mrStatinPreprintWeight`) of the weight is given to the preprinted result. However, the inclusion of the preprinted study did not have a substantial impact on the results (`r mrStatin$AD$estimate`) versus the single published study (`r estimate(0.59,0.25,1.39,"RR")`).

The other two preprints identified[@andrews2019; @zhu2017] reported on the effect of LDL-c on Alzheimer's disease using data from GLCC and IGAP consortia. As illustrated in Figure \@ref(fig:mrDuplication), these consortia were previously analysed in several published reports, and so these preprints did not add to the evidence base.

@@

Investigation of the publication status of the two preprints reporting studies not also identified by the main search after a two-year lag found that one had been subsequently published in late 2019.[@andrews2019; @andrews2021] The final preprint has not yet been published.[@so2017]

@@

<!-- TODO Check inclusion of Rockwood 2002:  -->
This review has presented a summary of the available evidence on the association between lipids, and treatments that modify lipids such as statins, and the subsequent risk of dementia. This discussion seeks to summarise the key findings in terms of literature sources and results as reported. A detailed comparison across the evidence sources, exposure measures and sources of bias reported here is presented as part of the triangulation exercise in Chapter \@ref(tri-heading).

@@

There was some indication of a protective effect of statins on all-cause and Alzheimer's disease dementia when looking at solely at observational studies. This finding was not supported by evidence from the two available RCTs, or by studies that emulated statin treatment using a genetic proxy, suggesting that these findings may be a result of heterogeneity in exposure (e.g. mid-life lipid lowering in non-randomised studies versus life-course lipid exposure in Mendelian randomisation studies and late-life lipid reduction in RCTs) or alternatively due to biases within the non-randomised studies.

@@

Across dementia outcomes, the majority of studies were non-randomised studies of lipids, or treatments that affect lipid levels such as statins. This distribution of evidence between analytical designs is to be expected. Randomised controlled trials of dementia are particularly challenging, as the long follow-up made necessary by the long latent period of the condition, makes trials logistically challenging and financial expensive. Similarly, Mendelian randomisation is a comparatively new study design (as illustrated in Figure \@ref(fig:typeByYear)), which appears in the literature in recent years, driven by the availability of summary genome wide association studies (GWAS) that form the basis of the two-sample Mendelian randomisation approach.

@@

A common theme was an absence of studies examining vascular dementia as an outcome, most noticeable when comparing the evidence base for statins in dementia/Alzheimer's (Figures \@ref(fig:obsStatinDementiaFig) & \@ref(fig:obsStatinADFig)) with that available for vascular dementia (Figure \@ref(fig:obsStatinVaDFig)). This is particularly interesting given that lipids and statins are strongly related to the prevention of vascular disease. A potential explanation for this observation may be publication bias or the "file-drawer effect",[@rosenthal1979] though there was weak evidence of a small-study effects for this outcome (of which publication bias is a potential cause).[@sterne2011] Similarly, only one Mendelian randomisation studies examined this outcome, primarily because of the absence (until recently) of vascular dementia GWAS which precludes a two-sample approach.

Of note, this review did not include the commonly cited PROSPER RCT, which examined the effect of pravastatin on CVD risk, reporting on cognitive outcomes as one of several secondary outcomes.[@shepherd2002] While widely cited in relation to the effect of statins on dementia risk and included in the Cochrane review of RCTs on this topic,[@mcguinness2016] the trial reported solely on the change in a range of cognitive measures (MMSE, Stroop test, Picture-Word Learning test and others) over follow-up. Though a useful indicator of general cognitive decline, it is not equivalent to a dementia diagnosis using recognised criteria as cognitive tests should feed into a broader diagnostic pathway (see Section \@ref(diagnostic-criteria)). As such, this trial did not meet the inclusion criteria for this review.

@@

Risk of bias across the individual results was generally quite high. The causes and expected impact of these biases on the results are discussed in more detail in Chapter \@ref(tri-heading). Of particular interest to this chapter, however, was the high risk of bias due to missing evidence observed for observational studies of lipid levels. In many cases, estimates were known to be missing from meta-analyses not at random due to preferential reporting of significant results observed in a number of analysis, leading to the high-risk judgement. These missing estimates were most commonly identified via analysis of conference abstract/final publication pairs.[@yamada2009conf; @yamada2009] In addition, some authors stated outright that non-significant results were not reported (e.g. "The other lipid variables not significantly associated with dementia and Alzheimer’s disease ... were not reported in the table.").[@ancelin2013] However, as all identified missing results are likely to be non-significant, they would not be expected to have a substantial impact on their respective meta-analysis (which provided weak evidence for an effect), other than increasing the precision of the summary estimate.

Finally in terms of generalisability, despite a large proportion of included studies being conducted in the Western world (Figure \@ref(fig:cohortLocations)), the applicability of the results to other populations is aided by the inclusion of studies which made use of data from the Taiwan health insurance database.

@@

While conducting this review, I identified several previous systematic reviews of this topic.[@chu2018; @yang2020; @muangpaisan2010; @poly2020;@kuzma2018stroke; @kuzma2018risk] However, this review is the first to use established domain based assessments tools (for example, the RoB 2 tool for randomized controlled trials)[@sterne2019] to assess the risk of bias in included studies. The majority of the highly cited reviews on this topic either do not formally consider risk of bias in the observational studies they include,[@chu2018;@power2015] or used a non-domain-based assessment tool (e.g. the Newcastle-Ottowa Scale).[@poly2020]

@@

Despite these differences in timescales and methodology, the duplication of work across reviews (including this review) is substantial. In retrospect, an alternative approach to conducting a further systematic review from scratch could have been employed. Known as an umbrella review, or review-of-reviews, these studies use other systematic reviews rather than primary studies as the unit of analysis.[@aromataris2015; @smith2011] This approach would have enabled more efficient identification of relevant primary studies to which the methods which sets this review apart from other published reviews could have been applied.

@@

todo

@@

Firstly, this study design is relatively new, particularly when compared to randomised trials or cohort studies. Figure \@ref(fig:typeByYear) demonstrates that Mendelian randomisation studies only begin to appear in the evidence base much later than NRSE/NRSI, likely due to the limited availability of large scale GWAS datasets needed for two-sample Mendelian randomisation analyses. As such, the process and tools for systematically assessing this study design are not as well developed. A key example of this is in the absence of validated search filters for Mendelian randomisation studies. This limitation is further complicated by the varying terminology used to describe the method, particularly in the early years of it's application, which led to me including general terms for instrumental variable analyses in my search.

Additionally, there is currently no widely used risk-of-bias assessment tool for Mendelian randomisation studies. While a recent commentary provided a checklist for interpreting Mendelian randomisation studies, this guide includes reporting items in their quality checklist.[@davies2018] While reporting quality is an important consideration, as reflected by the recent release of a STROBE_MR reporting guidelines,[@skrivankova2021] it is a separate consideration to internal validity  as discussed in Section \@ref(risk-of-bias). Similarly, a previous review of Mendelian randomisation studies used the Q-Genie tool which was validated to assess the quality of genetic association studies in meta-analysis.[@sohani2015] While this tool assesses the underlying GWAS used, it does not assess the additional methodological considerations of the Mendelian randomisation analysis itself. For this review, I used the best available author-devised tool, sourced from a recent review of systematic reviews of Mendelian randomisation studies.[@spiga2021]

>>>>

---
bibliography: bibliography/references.bib
csl: bibliography/nature.csl
output:
  bookdown::word_document2: 
      toc: false
      toc_depth: 3
      reference_docx: templates/word-styles-reference-01.docx
      number_sections: false 
  bookdown::pdf_document2:
    template: templates/brief_template.tex
  bookdown::html_document2: default
documentclass: book
---
```{block type='savequote', include=knitr::is_latex_output(),quote_author='(ref:sys-rev-results-quote)', echo = TRUE}
Hold  
```

@@

::: {.laybox data-latex=""}
## Lay summary {-}

Using the methods outlined in the previous chapter, here I present  the results of a systematic review of primary studies that have examined the relationship between the levels of blood lipids (such as cholesterol and triglycerides) and dementia outcomes. In addition, it examined the relationship between treatments that change blood lipid levels, such as statins, and dementia outcomes

My review included `r n_included` primary studies. I found that statins appear to reduce the risk of all-cause dementia and Alzheimer's disease, but were not associated with risk of vascular dementia. Lipid levels were not associated with any dementia outcome. However, problems with the methods used in some studies included in the review meant that I was less confident in their results. In addition, I found that several results were missing from my analysis as authors decided not to report them.

:::

<!-- TODO Complete! -->

@@

Following the methods outlined in the previous chapter, here I present the results of the comprehensive systematic review. The use of the evidence identified by the review in future chapters is discussed. Finally, I make recommendations for future methodological work around the inclusion of preprints and Mendelian randomisation studies in evidence synthesis exercises.

## Search results

@@

The database search identified `r num_to_text(23447)` records, of which `r comma(7338)` were duplicated records. Of the random sample of 500 records screened to ensure the accuracy of the study design filters, no eligible records were identified. As expected, many of those excluded by the filters were commentaries/educational articles, or described basic science studies. <!-- TODO include some example citations here -->

@@

The PRISMA flow diagram <!-- TODO CITATION NEEDED - cite PRIMSA flow paper here, if published --> presented in Figure \@ref(fig:prisma-flow-fig) illustrates the movement of articles through the review. To highlight the contribution of preprint archives to the review, the flow diagram delineates between those records captured through databases searches (presented on the left of the diagram) and those captured by the search tool described previously (presented in grey on the right of the diagram).

Common reasons for exclusion at the full text-stage included studies reporting on ineligible exposures (N = `r n_wrong_exposure` ; most commonly an ineligible lipid fraction) or outcomes (N = `r n_wrong_outcome`; e.g. change in cognitive scores), or usin an ineligible study design (N = `r n_study_design`).

@@

```{r prisma-flow-fig, echo = FALSE, results="asis", fig.pos="h", fig.cap='(ref:prisma-flow-cap)', out.width='100%', fig.scap='(ref:prisma-flow-scap)'}
knitr::include_graphics(file.path("figures/sys-rev/prismaflow.png"))
```

@@

For the assessment of the intra-/inter-rater reliability, the estimated values of Gwet's $AC1$ were interpreted against the categories presented in Table \@ref(tab:gwet-table). For the inter-rater reliability, agreement was "almost perfect" ($AC1$ = `r agreeInter_coeff[1]`, $kappa$ = `r agreeInter_coeff[2]`, Table \@ref(tab:agreeInter-table)). Similarly for intra-rater reliability, agreement was "almost perfect" ($AC1$ = `r agreeIntra_coeff[1]`, $kappa$ = `r agreeIntra_coeff[2]`, Table \@ref(tab:agreeIntra-table)). The discrepancy between the $AC1$ and $kappa$ coefficients illustrates the sensitivity of $kappa$ to imbalanced marginals, caused in this sample by a large imbalance  towards exclusion.[@feinstein1990]

@@

Those records which were excluded in the initial screening, but were included by the second reviewer (N =`r discrepancy_Inter`, Table \@ref(tab:agreeInter-table)) were investigated. This discrepancy between the two reviewers was explained in all cases by differing interpretations of the inclusion criteria, most commonly around the definition of cognitive decline versus dementia and of eligible lipids fractions.

@@

Following full-text screening, `r n_included` unique studies (described across `r n_reports_included` reports) met the criteria for inclusion in this review.`r get_all_citations()` Table \@ref(tab:studyCharacteristics-table) presents a summary of the characteristics of each study.

The majority of included studies described non-randomised analyses, with the only two included randomised controlled trials (the Heart Protection Study/British Heart Foundation trial[@heartprotectionstudycollaborativegroup2002] and the JUPITER trial[@ridker2008]) both reporting on the effect of statin use on incidence of all-cause dementia in older adults. A similarly small number of Mendelian randomisation studies were identified, several of which employed a two-sample approach using summary statistics from the same published genome wide association studies (GWAS), leading to complications in the synthesis (see Section \@ref(sys-rev-res-AD)).

@@

<!-- TODO Ensure all abbreviations at the end of this table are in the abbreviation section at the start. -->

@@

## Primary analyses

In the following sections, analyses are grouped first by outcome (all-cause dementia, Alzheimer's disease and vascular dementia) and then are further stratified by exposure category.

@@

(ref:statinsRCT-cap) Random-effects meta-analysis of randomised controlled trials examining statin statins on all-cause dementia. D1-D5 and O refer to the five risk-of-bias domains and single overall judgement in the RoB2 tool, while the risk of bias due to missing evidence is shown beside the overall summary diamond in each forest plot presented above.

@@

@@

<!----------------------------------------------------------------------------->
(ref:lipidFractionsDementia-cap) Random-effects meta-analysis of four lipid fractions (total cholesterol, HDL, LDL, and triglycerides) on all-cause dementia risk, standardised per 1-SD increase in the lipid fraction. Note that Tynkkyen _et al_[@tynkkynen2018] reported results from multiple cohorts within a single study and the results for each cohort are presented separately here.

@@

`r num_to_text(obsHyperchol$AD$n,T)` studies reported on the association of hypercholesterolemia with Alzheimer's disease and provided weak evidence for an effect (`r obsHyperchol$AD$estimate`, Figure \@ref(fig:obsHyperAD))

@@

<!----------------------------------------------------------------------------->
(ref:lipidFractionsAD-cap)  Random-effects meta-analysis of four lipid fractions (total cholesterol, HDL, LDL, and triglycerides) on Alzheimer's disease risk, standardised per 1-SD increase in the lipid fraction. Note that Tynkkyen _et al_[@tynkkynen2018] reported results from multiple cohorts within a single study and the results for each cohort are presented separately here.

@@

Finally, there were several identified Mendelian randomisation studies examining the effect of genetically lowered LDL-c on Alzheimer's disease risk (Figure \@ref(fig:mrDuplication)). However, all of these studies took a two-sample approach, making use of summary statistics from the GLGC and IGAP consortia. Due to this overlap, which would result in a falsely precise estimate caused by multiple counting of the same participants, no meta-analysis of these studies was performed. However, across each analysis using varying number of SNPS, no evidence for an effect was observed (Figure \@ref(fig:mrDuplication)).

@@

A single Mendelian randomisation analysis was identified that examining the effect of genetically lowered LDL-c levels via HMGCR inhibition on the risk of vascular dementia, which provided some evidence for an effect (`r mrStatin$VaD$estimate`).

<!-- TODO Might need to discuss this a bit further -->

@@

Few studies investigated the effect of individual lipid fractions. Only for total cholesterol was there greater than one result reported (Figure \@ref(fig:lipidFractionsVaD)), and a meta-analysis of these found weak evidence for an effect (`r n_effect(obsLipids$VaD_TC)`). A single study provided evidence on the other three fractions,[@yoshitake1995] and similar found minimal evidence of an effect LDL-c (`r n_effect(obsLipids$VaD_LDL,F)`), HDL-c (`r n_effect(obsLipids$VaD_HDL,F)`) or triglycerides (`r n_effect(obsLipids$VaD_TG,F)`).

@@

As shown in the forest plots in the above sections, the risk-of-bias assessments are presented alongside their corresponding numerical result. A more detailed discussion of the sources and directions of bias in each result is presented in Chapter \@ref(tri-heading), and so this section presents a brief summary of the biases observed in each study design.

For the two randomised controlled trials, both were judged to be at low risk of bias. In contrast, many of the non-randomised studies of statin use were at serious risk of bias primarily due to poor controlling for confounding, immortal time bias, and missing outcome data. Similarly, non-randomised studies of exposures suffered from incomplete adjustment for potentially important confounders, and concerns over the selection of the reported result from among several analyses (e.g. examination of lipids as a binary or continuous variable). Finally, bias was introduced into Mendelian randomisation studies via the potential for hoziontal pleiotropy and population stratification.[@davies2018]

Following best practice, any result judged to be at critical risk of bias should be excluded from any quantitative analyses.[@sterne2016] `r num_to_text(n_critical,T)` observational studies were excluded on this basis, predominantly due to a lack of adjustment for any potentially important confounders (i.e. the study reported unadjusted estimates).`r critical_citations`

## Additional analyses

@@

@@

There were `r n_dr_initial` studies initially considered eligible for the dose-response meta-analysis, as they appear to provide data on risk of dementia outcomes across several categories of lipid exposure. However, following data extraction, `r num_to_text(n_dr_excluded)` were excluded as they did not reported the relevant information needed (most commonly, the cut-off measures for each category).

Across the remaining `r num_to_text(n_dr_final)` studies, a sufficient number of results (n $\geqslant$ 3) were identified only for the total cholesterol-Alzheimer's, LDL-Alzheimer's, and total cholesterol-dementia strata. This analysis provided weak evidence for a non-linear effect of lipid levels on dementia outcomes, and Figure \@ref(fig:lipidsDoseResponse) illustrates this for the total cholesterol-Alzheimer's strata. Similar figures for the other analysed lipid-outcome strata are presented in Appendix \@ref(hold).

@@

### Heterogeneity and publication bias

Investigation of potential sources of heterogeneity was complicated by two factors. In the first instance, a minority of meta-analysis (4 out of 18, `r paste0(comma(4/18*100),"%")`) included more than 10 results, the recommended minimum required for meta-regression. Secondly, poor reporting of study characteristics of interest including education level and baseline cognitive ability precluded the use of several results in a meta-regression analysis.

<!-- TODO Finish this analysis -->

@@

@@

As show in the PRISMA flow diagram (Figure \@ref(fig:prisma-flow-fig)), the number of hits returned by preprint searching was not substantial (bioRxiv = 256, medRxiv = 0). From these hits, three preprinted reports of eligible studies were included in the review, of which two described unique studies not captured by the main search.[@andrews2019; @so2017]

One preprint (So _et al_[@so2017]) provided additional evidential value in a single meta-analysis (`r comma(1/18*100)`% of the 18 meta-analyses performed in this chapter). This meta-analysis of Mendelian randomisation studies examined the effects of lipid-lowering via HMGCR mutations on incidence of Alzheimer's disease (Figure \@ref(fig:mrStatinADFig)). To assess the evidence added to the meta-analysis through inclusion of the preprint, the data was re-analysed using a fixed-effect model. Examination of the weight assigned to each result in the analysis illustrates that a large proportion (`r mrStatinPreprintWeight`) of the weight is given to the preprinted result. However, the inclusion of the preprinted study did not have a substantial impact on the result of the meta-analysis (`r mrStatin$AD$estimate`) as compared to that reported by the single published study (`r estimate(0.59,0.25,1.39,"RR")`).

The other two preprints identified[@andrews2019; @zhu2017] reported on the effect of LDL-c on Alzheimer's disease using data from GLCC and IGAP consortia. As illustrated in Figure \@ref(fig:mrDuplication), these consortia were previously analysed in several published reports, and so these preprints did not add to new information to the evidence base.

@@

Investigation of the publication status of the two preprints reporting studies not also identified by the main search found that, allowing for a two-year lag, only one had been subsequently published.[@andrews2019; @andrews2021]

@@

<!-- TODO Check inclusion of Rockwood 2002:  -->
This review presented a summary of the available evidence on the association between lipids, and treatments that modify lipids such as statins, and the subsequent risk of dementia. This discussion seeks to summarise the key findings in terms of literature sources and results as reported. A detailed comparison across the evidence sources, exposure measures and sources of bias reported here is presented as part of the triangulation exercise in Chapter \@ref(tri-heading).

@@

There was some indication of a protective effect of statins on all-cause and Alzheimer's disease dementia when looking at solely at observational studies. This finding was not supported by evidence from the two available RCTs, or by studies that emulated statin treatment using a genetic proxy, suggesting that these findings may be a result of differing exposure windows (e.g. mid-life lipid lowering in non-randomised studies versus life-course lipid exposure in Mendelian randomisation studies and late-life lipid reduction in RCTs) or alternatively due to biases within the non-randomised studies.

@@

Across dementia outcomes, the majority of studies were non-randomised studies of lipids or lipid regulating agents. This distribution of evidence between analytical designs is to be expected. Randomised controlled trials of dementia are particularly challenging, as the follow-up made necessary by the long latent period of the condition makes trials logistically difficult and financial expensive. Similarly, Mendelian randomisation is a comparatively new study design, and studies employing it in relation to dementia outcomes only appear in the evidence base in recent years, as illustrated in Figure \@ref(fig:typeByYear). This recent increase is likely driven by the availability of summary genome wide association studies (GWAS) that form the basis of the two-sample Mendelian randomisation approach.

@@

A central finding of this review is the absence of studies examining vascular dementia as an outcome, most noticeable when comparing the evidence base for statins in dementia/Alzheimer's (Figures \@ref(fig:obsStatinDementiaFig) & \@ref(fig:obsStatinADFig)) with that available for vascular dementia (Figure \@ref(fig:obsStatinVaDFig)). This is particularly interesting given that lipids and statins are strongly related to the prevention of vascular disease. A potential explanation for this observation may be publication bias or the "file-drawer effect",[@rosenthal1979] though there was weak evidence of a small-study effects for this outcome (of which publication bias is one potential cause).[@sterne2011] Similarly, only one Mendelian randomisation study examined this outcome, likely due to the absence (until recently) of vascular dementia GWAS which precludes a two-sample approach.

Of note, this review did not include the commonly cited PROSPER RCT, which examined the effect of pravastatin on cardiovascular disease risk and reported on cognitive outcomes as one of several secondary outcomes.[@shepherd2002] While widely cited in relation to the effect of statins on dementia risk and included in the Cochrane review of RCTs on this topic,[@mcguinness2016] the trial reported solely on the change in a range of cognitive measures (MMSE, Stroop test, Picture-Word Learning test and others) over follow-up. Though a useful indicator of general cognitive decline, it is not equivalent to a dementia diagnosis using recognised criteria, as cognitive tests should feed into a broader diagnostic pathway (see Section \@ref(diagnostic-criteria)). As such, this trial did not meet the inclusion criteria for this review.

@@

Risk of bias across the individual results was generally high. The causes and expected impact of these biases on the results are discussed in more detail in Chapter \@ref(tri-heading). Of particular interest to this chapter, however, was the high risk of bias due to missing evidence observed for observational studies of lipid levels. In many cases, estimates are known to be missing not at random from meta-analyses due to preferential reporting of significant results, leading to high risk of bias due to missing evidence. These missing estimates were most commonly identified via analysis of conference abstract/final publication pairs.[@yamada2009conf; @yamada2009] In addition, some authors stated outright that non-significant results were not reported (e.g. "The other lipid variables not significantly associated with dementia and Alzheimer’s disease ... were not reported in the table."[@ancelin2013]). However, as all identified missing results are likely to be non-significant, they would not be expected to have a substantial impact on their respective meta-analysis (which provided weak evidence for an effect), other than increasing the precision of the summary estimate.

Finally in terms of generalisability, despite a large proportion of included studies being conducted in the Western world (Figure \@ref(fig:cohortLocations)), the applicability of the results to other populations is aided by the inclusion of several studies which made use of data from the Taiwan health insurance database.

@@

<!-- TODO Check none of these use a ROB tool -->

While conducting this review, I identified several previous systematic reviews of this topic.[@chu2018; @yang2020; @muangpaisan2010; @poly2020;@kuzma2018stroke; @kuzma2018risk] However, to my knowledge, this review is the first to use established domain based assessments tools (for example, the RoB 2 tool for randomized controlled trials)[@sterne2019] to assess the risk of bias in included studies. The majority of the highly cited reviews on this topic either do not formally consider risk of bias in the observational studies they include,[@chu2018;@power2015] or used a non-domain-based assessment tool (e.g. the Newcastle-Ottowa Scale).[@poly2020]

@@

Despite these differences in time scales and methodology, the duplication of work across systematic reviews (including this one) is substantial. In retrospect, an alternative approach to conducting a further systematic review from scratch, known as an umbrella review or review-of-reviews,[@aromataris2015; @smith2011] could have been employed. This study design uses other systematic reviews rather than primary studies as the unit of analysis, and would have enabled more efficient identification of relevant primary studies to which the methods which sets this review apart could have been applied.

@@

@@

Firstly, this study design is relatively new, particularly when compared to randomised trials or cohort studies. Figure \@ref(fig:typeByYear) demonstrates that Mendelian randomisation studies only begin to appear in the evidence base much later than NRSE/NRSI, likely due to the limited availability of large scale GWAS datasets needed for two-sample Mendelian randomisation analyses. As such, the process and tools for systematically assessing this study design are not as well developed. A key example of this is the absence of validated search filters for Mendelian randomisation studies. This limitation is further complicated by the varying terminology used to describe the method, particularly in the early years of its application, which led to me including general terms for instrumental variable analyses in my search.

Additionally, there is currently no widely used risk-of-bias assessment tool for Mendelian randomisation studies. While a recent commentary provided a checklist for interpreting Mendelian randomisation studies, this guide includes reporting items in their quality checklist.[@davies2018] Reporting quality is an important consideration for Mendelian randomisation studies, as reflected by the recent release of a STROBE MR reporting guidelines,[@skrivankova2021] but is a separate consideration to bias, as discussed in Section \@ref(risk-of-bias). Similarly, a previous review of Mendelian randomisation studies used the Q-Genie tool which was designed to assess the quality of GWAS included in a meta-analysis.[@sohani2015] While this tool assesses the underlying GWAS used, it does not address the additional methodological considerations of the Mendelian randomisation analysis itself. For this review, I used the best available author-devised tool, sourced from a recent review of systematic reviews of Mendelian randomisation studies.[@spiga2021]

As a further stumbling block, Mendelian randomisation lends itself to the analysis of multiple exposure-outcome comparisons as part of a single study, particularly when using a two-sample summary data design. This is particularly relevant to the consideration of bias due to missing evidence. As an example, through snowballing and other measures, I identified at least one relevant Mendelian randomisation study that had not been captured by the search strategy.[@larsson2017] This study examined lipid fractions as one of many risk factors for Alzheimer's disease, though the search would not have been expected to find it given the absence of any lipid-related keywords in the title or abstract. Studies examining multiple risk factors such as this can introduce bias into a systematic review, as it is commonly only those risk factors with statistically significant result that are reported in the abstract and so are captured by a systematic search. These studies are described as "unknown unknown's" in the context of the RoB-ME tool, and are considered to be particularly challenging (as opposed to an analysis that was insufficiently reported to be included in the statistical analysis, or the "known unknown's").
